---
title: "PetToLMANewData"
output: html_notebook
---


```{r}
require(BIEN)
require(plyr)
require(tidyr) 
require(dplyr)
require(mosaic)
require(stringr)
require(lme4)
require(magrittr)
require(data.table)
require(knitr)
require(kableExtra)
library(dataCompareR)
require(utils)
require(ggplot2)
require(merTools)
```

Load processed data
```{r}
royer_tax_new<-read.csv(file="royer_tax_new.csv")
fossildata<-read.csv(file="fossildata.csv")
##fix columns
fossildata <- fossildata[-c(1, 4)]
royer_tax_new <- royer_tax_new[-c(1, 3:6, 9:11)]
colnames(fossildata)[colnames(fossildata)=="Family"] <- "scrubbed_family"
colnames(fossildata)[colnames(fossildata)=="Genus"] <- "scrubbed_genus"
```

we now run the data through our model with a slope effect of log pet leaf area and intercept effect of scrubbed family.
```{r}
model<-lmer(log_lma~log_pet_leafarea + (1+log_pet_leafarea|scrubbed_family), data=royer_tax_new)
summary(model)
```

In order to compare our fossil data set to the living data set, we must omit predictions from fossildata that aren't present in royer_tax_new so all observations in the fossil data set are accounted for. 
```{r}
options(max.print = 1000)
af<-royer_tax_new$scrubbed_family
af
bf<-fossildata$scrubbed_family
bf

##whats in all_fossil_royer_pred that isn't in #royer_tax_new

missing<-bf[!(bf%in%af)]
write.csv(missing, file="missing.csv")

missing<-read.csv("missing.csv")
missingdata<-subset(missing, select=-X)
colnames(missingdata)[colnames(missingdata)=="x"]<-"scrubbed_family"

tallymissing<-
  missingdata%>%
  group_by(scrubbed_family)%>%
  tally()
print.data.frame(tallymissing)

fossildata<-fossildata[!fossildata$scrubbed_family %in% missingdata$scrubbed_family,]
```

Then, prediction intervals are created for LMA in fossildata based on the model. The dataset contains observations for expected LMA, but this gives prediction intervals for log LMA that give a more picture of what those values should be. 
```{r}
##Prediction Intervals 
PI<-predictInterval(merMod=model, newdata=fossildata, 
                    level=0.95, n.sims=1000,
                    stat="median", type="linear.prediction",
                    include.resid.var = TRUE)
PI
newfossilpredictions<-cbind(fossildata,PI)
ggplot(aes(x=log_pet_leafarea, y=fit, ymin=lwr, ymax=upr, color=fossildata$scrubbed_family), data=newfossilpredictions) +
  geom_point() +
  geom_linerange() +
  labs(color='Family')+
  xlab('log(Petiole Leaf Area)')+
  ylab('log(Leaf Mass Area)')+
  labs(title = 'Leaf Mass Predictions from Petiole Leaf Area with Family Information')
###save as PDF
ggsave("PettoLMAFigure.pdf")
```

Save fossil predictions and eliminate unnecessary columns
```{r}
write.csv(newfossilpredictions, file="PettoLMApredictionsfinal.csv")
```